---
title: "Statistical analisis"
author: "Alexandru Duda"
output:
  html_document:
    toc: true
    toc_depth: '2'
    df_print: paged
  html_notebook:
    toc: true
    toc_depth: 2
    number_sections: true
    theme: cerulean
    highlight: tango
  pdf_document:
    toc: true
    toc_depth: '2'
---
**About the study**

Two groups of participants will choose the grade of chondromalacia patellae according to the modified Outerbridge classification for 10 MRI images. The first group, hereafter the AI group, will receive a suggestion from artificial intelligence (AI) before choosing the grade. The second group, hereafter the non-AI group, will not receive this suggestion. The study will examine how the AI recommendation affects inter-observer variability.


Importing libraries.
```{r load-packages, message=FALSE, warning=FALSE}
library(pacman)
p_load(rio, dplyr, irr, boot)
```

The database is being imported. The file path must be entered manually and must correspond to the system being used.
```{r message=FALSE, warning=FALSE}
data <- import("C:/Users/dudaa/Documents/DataAnalisis/data.csv")
```

Overview of the database.
```{r}
glimpse(data)
```
```{r}
head(data)
```
The database contains 81 rows and 14 columns. Four columns contain categorical variables:
position – the participant's occupation. This can be student, resident, or physician.
year – applicable only to student participants. Refers to their current year of study.
country – refers to the country where they practice.
category – refers to the group they were assigned to. If TRUE, the participant is in the AI group; if FALSE, they are in the non-AI group.

The remaining 10 columns (from q1 to q10) contain the grades assigned by the participants to the images.

Graphs will be created for the categorical variables.
```{r}
barplot(table(data$position),
        names.arg = c("Resident", "Student"),
        main = "Role distribution",
        xlab = "Role",
        ylab = "Number of participants",
        ylim = c(0,100))

student_year <- table(data$year[ data$year != 0])
barplot(student_year,
        names.arg = c("Year 1", "Year 2", "Year 3", "Year 4", "Year 5", "Year 6"),
        main = "Year distribution for students",
        xlab = "Year",
        ylab = "Student number",
        ylim = c(0,25)
        )

barplot(table(data$country),
        names.arg = c("Germany", "R. of Moldova", "Romania", "Switzerland"),
        main = "Country distribution",
        xlab = "Country",
        ylab = "Number of participants",
        ylim = c(0,80)
)

barplot(table(data$category),
        names.arg = c("non-AI", "AI"),
        main = "Group distribution",
        #xlab = "Year of study",
        ylab = "Number of participants",
        ylim = c(0,50)
)
```
Conclusions that can be drawn after analyzing the graphs:
There is a relatively equal number of participants in the AI and non-AI groups.
Only Romania is sufficiently represented as a country.
Only the student group is sufficiently represented.

Two separate groups will be created based on the 'category' variable (the AI group and the non-AI group). The columns corresponding to the categorical variables will be removed. Only the columns containing the image classifications will be retained.
```{r}
# AI group
ai_data <- data[data$category == TRUE, !names(data) %in% c("country", "year", "position", "category")]
glimpse(ai_data)

# Non-AI group
non_ai_data <- data[data$category == FALSE, !names(data) %in% c("country", "year", "position", "category")]
glimpse(non_ai_data)
```
There are 41 participants in the AI group and 37 in the non-AI group.

To quantify inter-observer variability, Fleiss' Kappa will be used. The R language has a function for calculating this parameter (kappam.fleiss). This function returns several values in addition to the parameter itself, which are not needed. Below is a function that selects and returns only the Fleiss' Kappa value.
```{r}
compute_kappa <- function(ratings_matrix) {
  kappa_result <- kappam.fleiss(ratings_matrix)
  return(kappa_result$value)
}
```
The compute_kappa function requires a matrix where the images are rows and the participants' responses are columns—the inverse of the current format. A transposition of the initial matrix is necessary.

Fleiss' Kappa will be calculated for the AI group.
```{r}
ratings_ai <- t(as.matrix(ai_data))
kappa_ai <- compute_kappa(ratings_ai)
kappa_ai
```

Fleiss' Kappa will be calculated for the non-AI group.
```{r}
ratings_non_ai <- t(as.matrix(non_ai_data))
kappa_non_ai <- compute_kappa(ratings_non_ai)
kappa_non_ai
```
The difference in Fleiss' Kappa between the two groups will be calculated
```{r}
kappa_difference <- kappa_ai - kappa_non_ai
kappa_difference
```
Thus, Fleiss' Kappa in the AI group is higher than in the non-AI group. Moving forward, we will attempt to demonstrate that this result is statistically significant. To this end, the Bootstrap method will be used, as Fleiss' Kappa does not have a known sampling distribution.

To implement the bootstrap, we will use the boot function. it requires the following arguments:
data – the data being used
statistic – the function used to calculate the desired parameter
R – the number of repetitions (replicates)

The function used for the 'statistic' argument of the boot function receives the input data and an index vector to form a new group. For this new group, a new Fleiss' Kappa is calculated.
```{r}
boot_kappa <- function(data, indices) {
  sampled_data <- data[indices, ] # The new group is formed using the indices.
  
  # Calculating Fleiss Kappa
  kappa <- compute_kappa(t(sampled_data))
  return (kappa)
}
```

Bootstrap implementation for the AI group.
```{r}
# Bootstrap For AI group
boot_kappa1 <- boot(data = ai_data, statistic = boot_kappa, R = 20000)
# Confidence interval
boot.ci(boot_kappa1, conf = 0.99, type = c("perc"))
```

Bootstrap implementation for the non-AI group
```{r}
# Bootstrap for the non-AI
boot_kappa0 <- boot(data = non_ai_data, statistic = boot_kappa, R = 20000)
# Confidence interval
boot.ci(boot_kappa0, conf = 0.99, type = c("perc"))
```

Histograms of the values obtained from the two bootstrap procedures.
```{r message=FALSE, warning=FALSE}
hist(boot_kappa1$t,
     breaks = 50,
     col = "skyblue",
     main = "Histogram of Fleiss' Kappa (FK) values for the AI \n and non-AI groups obtained via the Bootstrap method", 
     xlab = "Fleiss Kappa",
     ylab = "Value frequency",
     ylim = c(0, 2000),
     xlim = c(0, 0.4)
     )
hist(boot_kappa0$t,
     breaks = 50,
     col = "orange",
     add = TRUE,
)
legend("topright",
       legend = c("FK in the non-AI group", "FK in the AI group"),
       fill = c("orange", "skyblue")
)
```
Visually, it can be observed that their intersection is minimal.

Histogram of the differences between the values of the two groups
```{r message=FALSE, warning=FALSE}
hist(boot_kappa1$t - boot_kappa0$t,
     breaks = 50,
     col = "forestgreen",
     main = "Histogram of Fleiss' Kappa differences \n between the two groups",
     xlab = "Fleiss Kappa AI  - Fleiss Kappa non-AI",
     ylab = "Value frequency",
     Xlim = c(0, 0.4),
     ylim = c(0,1500)
     )
```

The 99% confidence interval for this distribution will be calculated (one-tailed test; values between the 1st and 100th percentiles will be considered). The alternative hypothesis assumes that the AI intervention increases Fleiss' Kappa / decreases inter-observer variability.

k1 = Fleiss Kappa in the AI group


k0 = Fleiss Kappa in the non-AI group

Null hypothesis: mathematically, k1 - k0 <= 0. In words, Fleiss' Kappa in the AI group is less than or equal to that in the non-AI group.

Alternative hypothesis: mathematically, k1 - k0 > 0. In words, Fleiss' Kappa in the AI group is greater than in the non-AI group.
```{r}
quantile(boot_kappa1$t - boot_kappa0$t, probs = c(0.01, 1))
```

The 99% one-tailed confidence interval does not contain 0. In conclusion, the null hypothesis can be rejected.
